---
title: "Tooth Growth data analysis"
author: "Borye"
date: "Friday, November 07, 2014"
output: pdf_document
---

###Introduction###

This is a project for the statistical inference class hold by coursera. In this part of the project, I will analyze the ToothGrowth data in the R datasets packages

###1. Load the ToothGrowth data and perform some basic exploratory data analyses###

```{r}
library(lattice)
data(ToothGrowth)
str(ToothGrowth)
ToothGrowth$dose <- as.factor(ToothGrowth$dose)
bwplot(len ~ dose | supp, data = ToothGrowth)
```

###2. Provide a basic summary of the data###

```{r}
data(ToothGrowth)
len <- data.frame(ToothGrowth$len)
colnames(len) <- "len"
mean_len<- aggregate(len, list(supp = ToothGrowth$supp, dose = ToothGrowth$dose), mean)
sd_len <- aggregate(len, list(supp = ToothGrowth$supp, dose = ToothGrowth$dose), sd)
```

The mean of the dataset by supp is shownd below

```{r}
mean_len
```

The standard deviation of the dataset by supp is shownd below

```{r}
sd_len
```

###3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose###

Calculate the confidence interval of the tooth growth by different supp

```{r}
g1_supp <- mean_len[mean_len$supp == "OJ", ]$len
g2_supp <- mean_len[mean_len$supp == "VC", ]$len
t.test(g1_supp - g2_supp)
```

Calculate the confidence interval of the tooth growth by different dose

```{r}
g_dose_2 <- mean_len[mean_len$dose == 2.0, ]$len
g_dose_1 <- mean_len[mean_len$dose == 1.0, ]$len
g_dose_0.5 <- mean_len[mean_len$dose == 0.5, ]$len
t.test(g_dose_2 - g_dose_1)
t.test(g_dose_1 - g_dose_0.5)
t.test(g_dose_2 - g_dose_0.5)
```

###4. State your conclusions and the assumptions needed for your conclusions###

From the results we can observed that the **change in supp is the least efficient one**. The efficient of dose increasing is better then the change of supp. If we set the unit of increasing dose to 0.5. Then the **dose increasing from 0.5 to 1 is the most efficient one**. And the **confidence interval of the dose increasing from dose 0.5 to 1 is the narrowest**. 
