---
title: "Simulation Exercise Report"
author: "Borye"
date: "Friday, November 07, 2014"
output: pdf_document
---

###Introduction###

This is a project for the statistical inference class hold by coursera. In this part of the project, I will use simulation to explore inference and do some simple inferential data analysis.

The exponential distribution can be simulated in R with rexp(n, lambda) where lambda is the rate parameter. The mean of exponential distribution is 1/lambda and the standard deviation is also also 1/lambda. **Set lambda = 0.2** for all of the simulations. In this simulation, you will investigate the distribution of averages of 40 exponential(0.2)s. I will need to do a thousand or so simulated averages of 40 exponentials.

###Question 1, Show where the distribution is centered at and compare it to the theoretical center of the distribution###

First, I plot the original distribution of the rexp(40, 0.2)

```{r}
library(ggplot2)
p <- rexp(40, 0.2)
g = ggplot(data.frame(p = p), aes(x = p))
g = g + geom_histogram(color = "black", fill = "lightblue") + labs(title = "Distribution of rexp(40, 0.2)")
g
```

We could seen from this plot, that the distribution is similar as a exponential function. Next plot is the distribution of 1000 averages of 40 random rexp(40, 0.2) 

```{r}
mns = NULL
for (i in 1 : 100) mns = c(mns, mean(rexp(40, 0.2)))
vlines <- data.frame(xint = c(5, mean(mns)), grp = c("a", "b"))
g = ggplot(data.frame(mns), aes(x = mns))
g = g + geom_histogram(color = "black", fill = "lightblue") + geom_vline(data = vlines, aes(xintercept = xint, colour = grp), size = 1.5) + labs(title = "Distribution of 1000 averages of 40 random rexp(40, 0.2)")                                                                        
g
```

We could seen from this plot that the **red line shows the theoretical center of the distribution which is 5 (1/0.2)**. While the **blue line shows the center of the distribution**. We can seen from this plot that these two lines are very close to each other. The actual mean of the distributio is shown below. 

```{r}
mean(mns)
```

###Question 2, show how variable it is and compare it to the theoretical variance of the distribution###

The variance of the 1000 averages of 40 random rexp(40, 0.2) is shownd as below

```{r}
var(mns)
```

The theoretical variance of the distribution should be **(1/lambda)^2/n** which shownd as below

```{r}
(1/0.2)^2/40
```


###Question 3, show that the distribution is approximately normal###

From the plot, and the mean and varience of the 1000 averages of 40 random rexp(40, 0,2), we could believe that this distribution is approximately normal. 






