# Another Example
LI Bo

## Do not echo code

This is some text. Here is a code chunk

```{r simulation, echo=FALSE}
set.seed(1)
x <- rnorm(100)
mean(x)
```


## Hiding result

This is some text.

```{r simulation 1, echo=FALSE, results="hide"}
set.seed(2)
x <- rnorm(100)
mean(x)
```

## Inline text computations

```{r computetime, echo=FALSE}
time <- format(Sys.time(), "%a %b %d %X %Y")
rand <- rnorm(1)
```

The current time is `r time`. My favorite random number is `r rand`.

## Incorporating graphics

Let's first simulate some data.
```{r simulatedata, echo=TRUE}
x <- rnorm(100); y <- x + rnorm(100, sd = 0.5)
```

Here is a scatter plot of the data.
```{r scatterplot, fig.height=4}
par(mar = c(5, 4, 1, 1), las = 1)
plot(x, y, main = "My simulated Data")
```

## Making tables with xtable

```{r fitmodel}
library(datasets)
data(airquality)
fit <- lm(Ozone ~ Wind + Temp + Solar.R, data = airquality)
```

Here is a table of regression coefficients
```{r showtable, results="asis"}
library(xtable)
xt <- xtable(summary(fit))
print(xt, type = "html")
```

## Setting Global Options

```{r setoptions, echo=FALSE}
library(knitr)
opts_chunk$set(echo = FALSE, results = "hide")
```

First simulate data
```{r simulatedata 1, echo=TRUE}
x <- rnorm(100); y <- x + rnorm(100, sd = 0.5)
```

Here is a scatterplot of the data

```{r scaterplot 1, fig.height=4}
par(mar = c(5, 4, 1, 1), las = 1)
plot(x, y, main = "My simulated DATA")
```


